---
title: "Hearthstone Deck Data Analysis"
author: "Xinlin Zhang"
output: 
  html_document: 
    theme: cerulean
    toc: yes
    toc_depth: 4
---

Data: History of Hearthstone: 346,242 decks posted by players at Hearthpwn from 2013 to 2017.
<https://www.kaggle.com/romainvincent/history-of-hearthstone>

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(sqldf)
library(ggplot2)
library(lubridate)
data <- read.csv("~/Desktop/Blizzard/history-of-hearthstone/data.csv", stringsAsFactors=FALSE)
```

###Question: What are the performances of decks posted by players?
#####<span style="color:Black">(1) What is the number of decks by year?</span>
##### <span style="color:Black">(2) What's the relationship between dust cost and the popularity of decks? What are the ratings by deck class over time? </span>
#####<span style="color:Black">(3) What is the average dust cost by deck class?</span>
#####<span style="color:Black">(4) What are the most popular cards for different deck classes?</span>


###1. Key Findings 
##### <span style="color:Black">(1) The number of uploaded decks had steady growth from 2013 to 2016. However, the number varied greatly in different formats (standard and wild). 2015 had the lowest number of uploaded decks in the standard format but the highest number in the wild format.</span>
##### <span style="color:Black">(2) There is a weak relationship between the dust cost and the popularity of decks. In the standard format, class "Rogue" had the most ratings in 2013. The ratings for class "Warrior" outperformed other classes from mid-2014 to 2016. In the wild format, ratings decreased from 2014 to 2015 and then stabilized.</span>

##### <span style="color:Black">(3) Except "Warrior", deck classes cost more dust in the standard format than in the wild format. On average, class "Hunter" costs the least dust.</span>

##### <span style="color:Black">(4) Top 5 most popular cards for "Mage", "Priest" and "Rogue" in the standard format are similar to those in the wild format. </span>




### 2. Number of Decks by Year 
##### The number of uploaded decks increased over year. 
#####(Not considering the number in 2017 since data is only available before February 2017.)
```{r fig.height=2, fig.width=4, message=TRUE, warning=FALSE, paged.print=TRUE}
data$year<-year(as.POSIXct(data$date))
ggplot(sqldf('select count(deck_id) as number, year from data where year !=2017 group by year'),aes(year,number)) + geom_bar(stat = "identity")+ ylab("Number of Decks")+ xlab("Year")
```

##### In the standard format, the number of uploaded decks had a rocket increase in 2016.
```{r fig.height=2, fig.width=4, message=TRUE, warning=FALSE}
standard_deck <- data[data$deck_format=='S',]
standard_deck$year<-year(as.POSIXct(standard_deck$date))
sdeck<-sqldf('select count(deck_id) as number, year from standard_deck where year != 2017 group by year')
ggplot(sdeck,aes(year,number)) + geom_bar(stat = "identity")+ ylab("Number of Decks")+ xlab("Year")
```

##### In the wild format, the number of uploaded decks had a drastic increase in 2015 and decreased in 2016.
```{r fig.height=2, fig.width=4, warning=FALSE}
wild_deck <- data[data$deck_format=='W',]
wild_deck$year<-year(as.POSIXct(wild_deck$date))
wdeck<-sqldf('select count(deck_id) as number, year from wild_deck where year != 2017 group by year')
ggplot(wdeck,aes(year,number)) + geom_bar(stat = "identity")+ ylab("Number of Decks")+ xlab("Year")
```


### 3. Ratings by Deck Class Over Time
#####There is a weak relationship between the dust cost and the popularity of decks.
```{r fig.height=4, fig.width=4}
summary(lm(rating~craft_cost, data=standard_deck))
summary(lm(rating~craft_cost, data=wild_deck))
```

##### Standard format: Rogue had the most ratings in 2013 but went all the way to bottom in the following one year. Ratings of class 'Warrior' are higher than that of other classes between mid 2014 and 2016. After 2016, all deck classes equalized.

```{r fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
standard_deck <- data[data$deck_format=='S',]
standard_deck$date <- as.POSIXct(standard_deck$date)
ggplot(standard_deck, aes(date, rating, col=deck_class)) + geom_smooth(se=F) + ylab("Ratings")+ xlab("Year") + labs(col="Deck Class")
```

##### Wild format: Apart from Paladin and Druid, other deck classes had the highest ratings when released and decreased in the following one year. Ratings converged after 2015.

```{r fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
wild_deck <- data[data$deck_format=='W',]
wild_deck$date <- as.POSIXct(wild_deck $date)
ggplot(wild_deck, aes(date, rating, col=deck_class)) + geom_smooth(se=F) + ylab("Ratings")+ xlab("Year") + labs(col="Deck Class")
```



### 4. Average Dust Cost by Deck Class
##### Except class "Warrior", average dust cost is larger in standard format than in wild format. Mean craft cost for hunter is the smallest.

```{r fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
savg<-sqldf('select avg(craft_cost) as Standard, deck_class from standard_deck group by deck_class')

wavg<-sqldf('select avg(craft_cost) as Wild, deck_class from wild_deck group by deck_class')
avgcost<-merge(savg,wavg)
library(reshape2)
cost_class<-melt(avgcost,id.vars='deck_class',variable.name="Format",value.name="cost")
ggplot(cost_class,aes(deck_class,cost,fill=Format))+geom_bar(stat='identity',position='dodge')+
  ggtitle("Average dust cost by deck class")+ ylab("Dust Cost")+ xlab("Deck Class")

```


### 5. Most Popular Cards for Different Deck Class
#### (1) Standard Format

##### Druid
```{r message=TRUE, paged.print=TRUE}
s_druid<-subset(standard_deck, deck_class=="Druid")
s_druid_card<-as.data.frame(table(as.vector(as.matrix(s_druid[12:41]))))
colnames(s_druid_card)<-c("Card_number","Freq")
s_druid_card<-s_druid_card[order(s_druid_card[,2],decreasing=T),]
s_druid_card[1:5,]
```
##### Hunter
```{r echo=TRUE}
s_hunter<-subset(standard_deck, deck_class=="Hunter")
s_hunter_card<-as.data.frame(table(as.vector(as.matrix(s_hunter[12:41]))))
s_hunter_card<-s_hunter_card[order(s_hunter_card[,2],decreasing=T),]
s_hunter_card[1:5,]
```

##### Mage
```{r echo=TRUE}
s_mage<-subset(standard_deck, deck_class=="Mage")
s_mage_card<-as.data.frame(table(as.vector(as.matrix(s_mage[12:41]))))
s_mage_card<-s_mage_card[order(s_mage_card[,2],decreasing=T),]
s_mage_card[1:5,]
```

##### Paladin
```{r echo=TRUE}
s_paladin<-subset(standard_deck, deck_class=="Paladin")
s_paladin_card<-as.data.frame(table(as.vector(as.matrix(s_paladin[12:41]))))
s_paladin_card<-s_paladin_card[order(s_paladin_card[,2],decreasing=T),]
s_paladin_card[1:5,]
```

##### Priest
```{r echo=TRUE}
s_priest<-subset(standard_deck, deck_class=="Priest")
s_priest_card<-as.data.frame(table(as.vector(as.matrix(s_priest[12:41]))))
s_priest_card<-s_priest_card[order(s_priest_card[,2],decreasing=T),]
s_priest_card[1:5,]
```

##### Rogue
```{r echo=TRUE}
s_rogue<-subset(standard_deck, deck_class=="Rogue")
s_rogue_card<-as.data.frame(table(as.vector(as.matrix(s_rogue[12:41]))))
s_rogue_card<-s_rogue_card[order(s_rogue_card[,2],decreasing=T),]
s_rogue_card[1:5,]
```

##### Shaman
```{r echo=TRUE}
s_shaman<-subset(standard_deck, deck_class=="Shaman")
s_shaman_card<-as.data.frame(table(as.vector(as.matrix(s_shaman[12:41]))))
s_shaman_card<-s_shaman_card[order(s_shaman_card[,2],decreasing=T),]
s_shaman_card[1:5,]
```

##### Warlock
```{r echo=TRUE}
s_warlock<-subset(standard_deck, deck_class=="Warlock")
s_warlock_card<-as.data.frame(table(as.vector(as.matrix(s_warlock[12:41]))))
s_warlock_card<-s_warlock_card[order(s_warlock_card[,2],decreasing=T),]
s_warlock_card[1:5,]
```

##### Warrior
```{r echo=TRUE}
s_warrior<-subset(standard_deck, deck_class=="Warrior")
s_warrior_card<-as.data.frame(table(as.vector(as.matrix(s_warrior[12:41]))))
s_warrior_card<-s_warrior_card[order(s_warrior_card[,2],decreasing=T),]
s_warrior_card[1:5,]
```

##### Most popular cards in the standard format:
```{r echo=TRUE}
s_popular_cards<-cbind(s_druid_card[1:5,1],s_hunter_card[1:5,1],s_mage_card[1:5,1],s_paladin_card[1:5,1],s_priest_card[1:5,1],s_rogue_card[1:5,1],s_shaman_card[1:5,1],s_warlock_card[1:5,1],s_warrior_card[1:5,1])
colnames(s_popular_cards)<-c('Druid','Hunter','Mage','Paladin','Priest','Rogue','Shaman','Warlock','Warrior')
s_popular_cards
```

#### (2) Wild Format
##### Druid
```{r message=TRUE, paged.print=TRUE}
w_druid<-subset(wild_deck, deck_class=="Druid")
w_druid_card<-as.data.frame(table(as.vector(as.matrix(w_druid[12:41]))))
colnames(w_druid_card)<-c("Card_number","Freq")
w_druid_card<-w_druid_card[order(w_druid_card[,2],decreasing=T),]
w_druid_card[1:5,]

```

##### Hunter
```{r echo=TRUE}
w_hunter<-subset(wild_deck, deck_class=="Hunter")
w_hunter_card<-as.data.frame(table(as.vector(as.matrix(w_hunter[12:41]))))
w_hunter_card<-w_hunter_card[order(w_hunter_card[,2],decreasing=T),]
w_hunter_card[1:5,]
```

##### Mage
```{r echo=TRUE}
w_mage<-subset(wild_deck, deck_class=="Mage")
w_mage_card<-as.data.frame(table(as.vector(as.matrix(w_mage[12:41]))))
w_mage_card<-w_mage_card[order(w_mage_card[,2],decreasing=T),]
w_mage_card[1:5,]
```

##### Paladin
```{r echo=TRUE}
w_paladin<-subset(wild_deck, deck_class=="Paladin")
w_paladin_card<-as.data.frame(table(as.vector(as.matrix(w_paladin[12:41]))))
w_paladin_card<-w_paladin_card[order(w_paladin_card[,2],decreasing=T),]
w_paladin_card[1:5,]
```
##### Priest
```{r echo=TRUE}
w_priest<-subset(wild_deck, deck_class=="Priest")
w_priest_card<-as.data.frame(table(as.vector(as.matrix(w_priest[12:41]))))
w_priest_card<-w_priest_card[order(w_priest_card[,2],decreasing=T),]
w_priest_card[1:5,]
```

##### Rogue
```{r echo=TRUE}
w_rogue<-subset(wild_deck, deck_class=="Rogue")
w_rogue_card<-as.data.frame(table(as.vector(as.matrix(w_rogue[12:41]))))
w_rogue_card<-w_rogue_card[order(w_rogue_card[,2],decreasing=T),]
w_rogue_card[1:5,]
```

##### Shaman
```{r echo=TRUE}
w_shaman<-subset(wild_deck, deck_class=="Shaman")
w_shaman_card<-as.data.frame(table(as.vector(as.matrix(w_shaman[12:41]))))
w_shaman_card<-w_shaman_card[order(w_shaman_card[,2],decreasing=T),]
w_shaman_card[1:5,]
```

##### Warlock
```{r echo=TRUE}
w_warlock<-subset(wild_deck, deck_class=="Warlock")
w_warlock_card<-as.data.frame(table(as.vector(as.matrix(w_warlock[12:41]))))
w_warlock_card<-w_warlock_card[order(w_warlock_card[,2],decreasing=T),]
w_warlock_card[1:5,]
```

##### Warrior
```{r echo=TRUE}
w_warrior<-subset(wild_deck, deck_class=="Warrior")
w_warrior_card<-as.data.frame(table(as.vector(as.matrix(w_warrior[12:41]))))
w_warrior_card<-w_warrior_card[order(w_warrior_card[,2],decreasing=T),]
w_warrior_card[1:5,]
```

##### Most popular cards in the wild format:
```{r echo=TRUE}
w_popular_cards<-cbind(w_druid_card[1:5,1],w_hunter_card[1:5,1],w_mage_card[1:5,1],w_paladin_card[1:5,1],w_priest_card[1:5,1],w_rogue_card[1:5,1],w_shaman_card[1:5,1],w_warlock_card[1:5,1],w_warrior_card[1:5,1])
colnames(w_popular_cards)<-c('Druid','Hunter','Mage','Paladin','Priest','Rogue','Shaman','Warlock','Warrior')
w_popular_cards
```

###Appendix: Data Description
#####Actual Hearthstone deck records. Each record features :
```{r fig.height=8, fig.width=10}
library(readxl)
Description <- read_excel("~/Desktop/Description.xlsx")
Description
```




